---
title: "Example Weekly Report"
subtitle:
- <h1><u>CONFIDENTIAL-DO NOT DISSEMINATE</u></h1>
author: "Sam Hamilton"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document:
    toc: true
    toc_depth: 2
---


```{r, echo=FALSE}
# Purpose: validate incoming data collected for XXX study

# Additions to make:
  # Add table of birthweights
  # Estimate the study finish date by extrapolating the enrollment rate
```

## Automating the data export process
```{r setup, include=FALSE}
# set date
date <- format(Sys.Date(), "%d%b%Y")

# configure directories, load libraries and base functions
source(paste0(here::here(), "/monitoring/example_report/code/functions/config_example.R"))

# source in functions-refer to these scripts for automated data export
source(paste0(code_dir, "export_data_example.R"))
source(paste0(code_dir, "api_example.R"))

#read in metadata for each api call
api_meta <- fread(paste0(tab_dir, "api_meta.csv"))

#Function arguments that won't change
download_folder <- paste0(data_dir)
data_loc <- paste0(data_dir, "export_", date)
headquarters_address <- "https://gamin.mysurvey.solutions"

#Run api--This will take some time...
if(!dir.exists(data_loc)){
  export.data(download_folder, data_loc, headquarters_address, api_meta)
}
```

## Data cleaning/processing
```{r clean and read data, include=F}
source(paste0(code_dir, "read_clean_data_example.R"), local = T)
```

## Data monitoring
### Data validation
#### Missing/inappropriately structured data
```{r, include=F}
#check for missing values across all columns
map(dt.anthro, ~sum(is.na(.)))

#check for missing values in childsex
na_childSex <- sum(is.na(dt.anthro$childSex))
```

There are `r na_childSex` rows missing child sex value.

#### Reporting inconsistencies
```{r, include=F}

```



### Data presentation (DSMC-style)
#### Enrollment and baseline characteristics
```{r enrollment counts, include=F}
#create screened indicator
dt.eligibility %<>%
  mutate(screened = if_else(!duplicated(childID) | is.na(childID), 1, 0))

#create eligible indicator-based on eligibility criteria(age, oral feeding, not moving away, no allergy, lives in area)
dt.eligibility %<>%
  mutate(eligible = if_else(ageInDays>7 &
         ageInDays<1826 &
         feedOrally == 1 &
         available6months == 1 & 
         allergy == 2 &
         otherStudy == 2 &
         primaryResidence==1 &
         screened == 1, 1, 0))

#create enrolled indicator
dt.eligibility %<>%
  mutate(enrolled = if_else(eligible == 1 & consent == 1 & !is.na(childID), 1, 0))
         
#create refused indicator
dt.eligibility %<>%
  mutate(refused = if_else(consent == 2 | (is.na(consent) & eligible == 1), 1, 0))

#tabulate the indicators
temp <- dt.eligibility %>%
  summarise(Screened = sum(screened),
            Eligible = sum(eligible),
            Refused = sum(refused),
            Enrolled = sum(enrolled))

#tabulate the indicators for past 7 days
temp2 <- dt.eligibility %>%
  filter(Sys.Date() - startTime < 7) %>%
  summarise(Screened = sum(screened),
            Eligible = sum(eligible),
            Refused = sum(refused),
            Enrolled = sum(enrolled))


#create tables for overall and past 7 days
temp.table <- kable(temp, "latex", align = "c", linesep = '') %>%
  row_spec(0,bold=TRUE)
temp.table2 <- kable(temp2, "latex", align = "c", linesep = '') %>%
  row_spec(0,bold=TRUE)
```

```{r Cumulative enrollment}
#daily enrollments over time
total_perday <- dt.eligibility %>%
  group_by(startTime) %>%
  summarise(Screened = sum(screened), 
            Eligible = sum(eligible),
            Enrolled = sum(enrolled))
  
#Melt to long for ggplot
total_perday %<>%
  gather(key = "Key", value = "Value", Screened, Eligible, Enrolled) %>%
  group_by(Key) %>%
  mutate(total=cumsum(Value)) %>%
  arrange(Key, startTime)

#plot
total_perday$Step <- factor(total_perday$Key, levels=c("Screened", "Eligible", "Enrolled"))
ggplot(total_perday, aes(x = startTime, y = as.integer(total), color = Step)) +  
  scale_color_brewer(palette="Set1", name="") +
  geom_line(alpha=.5) +
  geom_point() +
  xlab("Date") +
  ylab("Total participants") +
  scale_y_continuous(breaks= pretty_breaks()) +
  scale_x_date(labels = date_format("%m/%d"), breaks = date_breaks("1 day")) +
  theme(legend.title = element_blank()) +
  #scale_color_discrete(name="") +
  theme_bw()
```

#### Treatment
```{r treatment letter validation, include=F}
#read in table that links each ID with respective treatment letter (randomized)
dt_letters <- read_csv(paste0(tab_dir, "treatment_letters.csv"))

#merge with treatment letters
wrong_random <- dt.treatment %>%
  left_join(dt_letters, by="childID") %>%
  #the treatment letter does not match what's in database
  tally(TL != assignment) %>% as.numeric()
```

```{r treatment coverage table, include=F}
#treatment coverage
enrolled <- dt.eligibility %>%
  summarise(Enrolled = sum(enrolled))

#how many kids treated
treat_cov <- dt.treatment %>%
  filter(isAdministered==1) %>%
  summarise(Treated = length(unique(childID))) %>%
  bind_cols(enrolled)

#calculate coverage percent
treat_cov %<>%
  mutate(Coverage = percent(Treated/Enrolled, 1))

#table
kable(treat_cov, "latex", booktabs = T, align = "c", linesep = '') %>%
  row_spec(c(0, bold_rows),bold=TRUE)

```

#### Primary/secondary outcomes (***masked!***)
```{r primary outcome, echo=FALSE}
#number of deaths by city and overall
n_died6mo <- dt.followup[vitalStatus == 2 & timepoint<=180, .(Deaths = .N), by=region_name]
if(nrow(n_died6mo) == 1) n_died6mo <- rbind(n_died6mo, 
                                            data.table(region_name="Boucle du Mouhoun",
                                                       Deaths = 0))
n_died6mo.tot <- dt.followup[vitalStatus == 2 & timepoint<=180, 
                             .(Deaths = .N, region_name="Overall")]
n_died6mo <- rbind(n_died6mo, n_died6mo.tot)

#person time by city and overall
dt.persont.tab <- dt.persont[,.(person_time = round(sum(person_days, na.rm = T)/365,2)),
                             by="region_name"]
dt.persont.tot <- dt.persont[,.(person_time = round(sum(person_days, na.rm = T)/365,2),
                                region_name="Overall")]
dt.persont.tab <- rbind(dt.persont.tab, dt.persont.tot)

#merge together deaths and person time to calculate rate
death6rate <- merge(n_died6mo, dt.persont.tab, by="region_name", all.y = T)
death6rate[is.na(Deaths), Deaths := 0]
death6rate[,mortality_rate := round((Deaths/person_time)*1000, 2)]
bold_rows <- which(death6rate[,region_name]=="Overall")
kable(death6rate, col.names = c("Region", "Deaths", "Person-years", 
                                "Mortality (per 1000 py)")) %>%
    row_spec(c(0, bold_rows),bold=TRUE)
```

```{r growth curves, echo=FALSE}
#labels for SD in plot
sd_labels <- factor(c("-3 SD", "-2 SD", "-1 SD", "Mean", "+1 SD", "+2 SD", "+3 SD"),
                    levels = c("-3 SD", "-2 SD", "-1 SD", "Mean", "+1 SD", "+2 SD", "+3 SD"))
p_frmSD <- c(.2, 2.3, 15.9, 50, 84.1, 97.7, 99.8)

#height-require growthstandards package
chartdat_w <- data.frame(SD = rep(sd_labels, each = 21),
                       Day = rep(8:28, times = 7),
                       weight_male= unlist(map(p_frmSD, 
                                               function(p) 
                                                 who_centile2htcm(8:28, p, sex="Male"))),
                       weight_female = unlist(map(p_frmSD, 
                                                  function(p) 
                                                    who_centile2htcm(8:28, p, sex="Female"))))

#format baseline and follow-up form for binding to each other
temp <- dt.baseline[,.(childID, childBirthdate, weight, childSex=q_2_1_1,
                       length, days, startDate)]
temp[, form := "Baseline"]
temp2 <- dt.followup[timepoint==180, .(childID, childSex, length, weight,
                                       form, startDate = as.Date(startTime))]
temp2[, form := "6-month"]

#merge on birthdate from baseline b/c it's likely more accurate than follow-up
temp2 <- merge(temp2, dt.baseline[,.(childID, childBirthdate)], by="childID")
temp2[, days := difftime(startDate, childBirthdate, units = "days")]
dt.plot <- rbind(temp, temp2)
dt.plot$form <- factor(dt.plot$form, levels=c("Baseline", "6-month"))

#plot males
oldest <- dt.plot[,as.numeric(max(days))]
p.male.hfa6 <- ggplot(dt.plot[childID %in% unique(temp2$childID) & childSex==1],aes(days, length)) +
  geom_point(aes(shape=form)) +
  geom_line(data=table.hfa.boys[Day<=oldest], aes(Day, length, color=SD)) +
  theme_bw() +
  ggtitle("Males") +
  xlab("Age (days)") +
  ylab("Height (cm)") +
  scale_color_discrete(name="WHO Standard") +
  scale_shape_discrete(name="Timepoint") +
  #ylim(min(dt.plot[length>0, length]), max(dt.plot[,length])) +
  guides(color = guide_legend(reverse = TRUE))
```

#### Follow-up/Adverse events

```{r followup verification, echo=FALSE}
#IDs of kids that are at least 7 days post treatment and have 7, 14, or 21 day follow-up
died <- dt.followup[vitalStatus==2, unique(childID)]
dt.eligibles <- dt.treatment[startDate + 21 <= Sys.Date() & 
                               childID %in% setdiff(childID, died), .(childID, region_name)]
followupID <- dt.followup[timepoint %in% c(7,14,21), unique(childID)]

#create variable that indicates if ID has had 7, 14, or 21 day followup
dt.eligibles[childID %in% followupID, follow_21 := 1]
dt.eligibles[is.na(follow_21), follow_21 := 0]

#collapse for table
tab <- dt.eligibles[,.(followup = percent(mean(follow_21))), by="region_name"]
tot <- dt.eligibles[,.(followup = percent(mean(follow_21)), region_name = "Overall")]
tab <- rbind(tab, tot)
bold_rows <- which(tab[,region_name]=="Overall")
kable(tab, col.names = c("Region", "Follow-up")) %>%
  row_spec(c(0, bold_rows),bold=TRUE)
```

```{r adverse events, echo=F}
#adverse event counts by region
dt.adverse <- dt.followup %>%
select(childID, region_name, timepoint, q_5_6__1, q_5_6__2, q_5_6__3, 
       q_5_6__4, q_5_6__5, q_5_6__6) %>%
gather(key="key", value = "value",
       q_5_6__1, q_5_6__2, q_5_6__3, q_5_6__4, q_5_6__5, q_5_6__6) %>%
filter(value==1 | value==0) %>%
group_by(key, region_name) %>%
summarise(total=sum(value)) %>%
spread(key = "key", value = "total") %>%
rename(Region=region_name, Fever=q_5_6__1, Diarrhea=q_5_6__2, Vomiting=q_5_6__3, 
       Abdominal=q_5_6__4, Rash=q_5_6__5, Constipation=q_5_6__6)

#add on overall row
dt.adverse <- bind_rows(dt.adverse,
          dt.adverse %>%
            summarise(Region="Overall", Fever=sum(Fever), 
                      Diarrhea=sum(Diarrhea), Vomiting=sum(Vomiting),
                      Abdominal=sum(Abdominal), Rash=sum(Rash),
                      Constipation=sum(Constipation))) %>%
  rename("Abdominal Pain"=Abdominal) %>% as.data.table()

#make kable
bold_rows <- which(dt.adverse[,Region]=="Overall")
kable(dt.adverse, "latex", align = "c") %>%
  row_spec(c(0, bold_rows),bold=TRUE) %>%
  collapse_rows(1, row_group_label_position = 'identity')
```

### Appendix
```{r other ae list, echo=FALSE}
#restrict data to hospital visits for reasons "other"
otherhos_visits <- dt.followup[reasonHosVisit__5 == 1, 
                               as.character(unique(specifyHosVisit))]
other_visits <- dt.followup[reasonVisit__5 == 1, 
                            as.character(unique(specifyVisit))]
other_visits <- c(other_visits, otherhos_visits)
sort(unique(other_visits))
```